### 2023-03-27 response-to-audit-report.md

```console
commit fe00af7d622bbd17c3327363d4650ee544d1fff7
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 11:25:09 2023 -0700

    PLT-4169 Added text to specification on how to mint role tokens.

    This addresses the audit comment:

    > File marlowe-cardano-specification.md, Section Monetary Policy
    > for Role Tokens The minting policy is not specified, but a
    > reference needs to be offered to explain how to create one.
```

@@

### 2.5.3 Argument for Contract in `txInfoData` not specified *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Section `Types`** 
> 
> The argument by which the `Contract` in the `txInfoData` list corresponds to the given hash needs to be made explicit.

TODO: @bwbush will add text here.

```console
commit b6366efc9f8df38bfe9232f81e6caf2c7f11ffa8
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 11:31:11 2023 -0700

    PLT-4169 Made explicit the relationship between continuation and its hash.

    This addresses the audit comment:

    > File marlowe-cardano-specification.md, Section Types The argument
    > by which the Contract in the txInfoData list corresponds to the
    > given hash needs to be made explicit.
```

@@

### 2.5.4 Merkleization section not detailed enough *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Section `Merkleization`**
> 
> This section is too terse. It needs to explain what Merkleization is, and to motivate why it is needed.
> 
> When explaining how it works, it needs to make explicit that only the `Case` type is modified, and that in the semantics, only the `Input` type is modified. It needs to explain why the `Input` type needs to carry a hash and a contract, and why the evaluation of the contract is changed as described.

TODO: @bwbush will add text here.

```console
commit a9d4043230cb4f240fab2d816bed6a3cab4f5053
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 11:44:32 2023 -0700

    PLT-4169 Elaborated explanation of and motivation for merkleization.

    This addresses the audit comment:

    > File marlowe-cardano-specification.md, Section Merkleization
    > This section is too terse. It needs to explain what Merkleization
    > is, and to motivate why it is needed. When explaining how it works,
    > it needs to make explicit that only the Case type is modified, and
    > that in the semantics, only the Input type is modified. It needs
    > to explain why the Input type needs to carry a hash and a contract,
    > and why the evaluation of the contract is changed as described.
```

@@

### 2.5.5 Unnecessary constraint *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, `Constraint 12. Merkleized continuations`** 
> 
> This constraint is unnecessary to have in the Marlowe validator, since the construction of the arguments for evaluation of the Marlowe contract would fail. However, it would be useful to have it appear in the specification for users to be aware of it when crafting transactions. A note to motivate the presence of the constraint could be helpful.

TODO: @bwbush will add text here.

```console
commit fda1d3e862fc8f8c31ff479fbd338226e661eb0b
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 11:50:20 2023 -0700

    PLT-4169 Clarified nature of Constraint 12 in specification.

    This addresses the auditor comment:

    > File marlowe-cardano-specification.md, Constraint 12. Merkleized
    > continuations This constraint is unnecessary to have in the
    > Marlowe validator, since the construction of the arguments for
    > evaluation of the Marlowe contract would fail. However, it would
    > be useful to have it appear in the specification, for users to
    > be made aware of it when crafting transactions. A note to motivate
    > the presence of the constraint could be helpful.
```

@@

### 2.5.6 Asymmetry between role and wallet payouts *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, `Constraint 15`, **
> 
> The marlowe validator allows multiple outputs to be paid to a wallet, but it demands that a single output exists when paying to a role instead. The motivation to use different approaches needs to be documented. This is implemented in `Scripts.hs` at line 371, in function `payoutConstraints`.

TODO: @bwbush will add text here.

```console
commit 4e81470de90ddd6858c3ad8aace96e90df5a86ae
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Wed Mar 1 17:40:38 2023 -0700

    SCP-5128 Commented on rationale for allowing multiple outputs to an address.
```

@@

### 2.5.7 Incorrect description of `rolePayoutValidator` *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Section `Plutus Validator for Marlowe Payouts`** 
> 
> The description of the Marlowe payout validator in the specification states that it is parameterized by the currency symbol. However, this is not correct as the validator is unparameterized; rather, the datum type of the validator includes the currency symbol (as well as token name). The description should be modified to reflect this.

TODO: @bwbush will add text here.

```console
commit 1d16de1f1b7a167dc4eb4a057a3df2c24d0194c9
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 11:12:36 2023 -0700

    PLT-4169 Fixed example call to `rolePayoutValidator`.

    This addresses the following audit finding:

    > File marlowe-cardano-specification.md, Section Plutus Validator
    > for Marlowe Payouts The type signature of rolePayoutValidator
    > is missing the datum argument. It should be: rolePayoutValidator
    > :: CurrencySymbol -> TokenName -> () -> ScriptContext -> Bool
```

@@

### 2.5.8 Unspecified initial state *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Section `Life Cycle of a Marlowe Contract`** 
> 
> The specification should say what the initial state of a Marlowe contract should be. In particular, creating a contract requires giving the minimum Ada to some account in the Marlowe state. Otherwise, Constraint 5 will reject the transactions that try to spend the output.

TODO: @bwbush will add text here.

```console
commit e150e30867dfa5466639f46bf39656aee2888e38
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 12:34:38 2023 -0700

    PLT-4169 Elaborated in specification on the required initial state.

    This addresses the audit comment:

    > File marlowe-cardano-specification.md, Section Life Cycle of a
    > Marlowe Contract The specification should say what the initial
    > state of a Marlowe contract should be. In particular, creating
    > a contract requires giving the minimum Ada to some account in
    > the Marlowe state. Otherwise, Constraint 5 will reject the
    > transactions that try to spend the output.
```

@@

### 2.5.9 Unspecified behavior when multiple cases can apply *(Severity: Low)*

> **File `Semantics.hs`, Function `applyCases`, line *597***
> 
> If multiple cases in a case list can apply, the first one is taken. This behavior should be better communicated in the specification.

TODO: @bwbush will add text here.

```console
commit 08d6f34ac91da7aa34e7414c5bae0e33e9235d74
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 12:06:10 2023 -0700

    PLT-4169 Improved documentation of `applyCases`.

    This addresses the audit comment:

    > File Semantics.hs, Function applyCases, line 597 If multiple
    > cases in a case list can apply, the first one is taken. This
    > behavior should be better communicated in the specification.
```

@@

```console
```

### 2.6.3 Variable names differ from Isabelle code *(Severity: Low)*

> **File `Semantics.hs`, Several functions ``** 
> 
> Different variable names in Isabelle and Haskell make comparison harder. It is less of an issue when only one variable has been renamed in a function, but multiple variable renames require carefully mapping between the names to avoid confusion. For example, the code of `reduceContractStep` in line 482 (`Pay` case) is hard to compare.
> 
> Other name changes include:
> -   Line 456, function `updateMoneyInAccount` uses variable `money` where the Isabelle code uses `amount` and omits naming the last parameter.
> -   Line 473, function `giveMoneyToPay` uses variables `amount` and `accounts` instead of `money` and `accountsV` as in the Isabelle code.
> -   Line 541, function `reductionLoop` uses variable `con` instead of `ncontract`.
> -   Line 300, the data type `TransactionInput` corresponds to the type `Transaction` in the Isabelle code.
> -   Line 313, the data type `TransactionOutput` is isomorphic but not identical to the homonymous data type in the Isabelle code.
> -   Line 439, function `refundOne` uses a variable `balance` where the Isabelle code uses `money`.
> -   Line 463, function `addMoneyToAccount` uses variable `accounts` where the Isabelle code uses `accountsV`.

TODO: @bwbush will add text here.

```console
commit 1c96edfe7e22ecc499ed75dd175490a4f2427bdc
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 13:24:06 2023 -0700

    PLT-4169 Renamed functions, per audit recommendations.
```

@@

TODO: @bwbush will add text here.

```console
```

### 2.6.9 `computeTransaction` differs from the Isabelle implementation *(Severity: Low)*

> **Helper function `evalValue, evalObservation`, lines *391 (Semantics.hs), 34 (Semantics.thy)***
> 
> `evalValue` and `evalObservation` differ from the Isabelle implementation in the introduction of auxiliary variables to abbreviate the recursive calls. The comparison would be simpler if both definitions were consolidated.

> **Helper function `evalValue`, lines *395 (Semantics.hs), 35 (Semantics.thy)***
> 
> The Isabelle implementation should use the helper function `moneyInAccount` instead of inlining its definition, so as to maintain consistency with the Haskell implementation.

> **Helper function `applyCases`, lines *596 (Semantics.hs), 498 (Semantics.thy)***
> 
> The structure of function `applyCases` differs between the Haskell and Isabelle files. Specifically, the Haskell implementation has an additional function `applyAction` where the Isabelle implementation does not. A comment motivating the discrepancy would be needed. This is likely due to the lack of Merkleization in the Isabelle implementation.

TODO: @bwbush will add text here.

```console
commit d0a3834ae8484bb6c7d0ecd3aeec60871b3fd165
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 13:00:28 2023 -0700

    PLT-4169 Added comment about merkleization to `applyCases`.

    This addresses the audit comment:

    > Helper function applyCases, lines 596 (Semantics.hs), 498
    > (Semantics.thy) The structure of function applyCases differs
    > between the Haskell and Isabelle files. Specifically, the
    > Haskell implementation has an additional function applyAction
    > where the Isabelle implementation does not. A comment
    > motivating the discrepancy would be needed. This is likely
    > due to the lack of Merkleization in the Isabelle implementation.
```

> **Helper function `convertReduceWarnings`, lines *617 (Semantics.hs), 537 (Semantics.thy)***
> 
> The Haskell function is implemented using `foldr`, while the Isabelle function uses explicit recursion, making a one-to-one comparison less obvious. If there is a reason for this discrepancy, such as `foldr` yielding some optimizations, this should be outlined in a comment.

TODO: @bwbush will add text here.

```console
commit d14776101c22eac7fbc67893020f37cbc2a9e339
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 13:02:40 2023 -0700

    PLT-4169 Added note on use of `foldr` in semantics.

    This addresses the audit comment:

    > Helper function convertReduceWarnings, lines 617 (Semantics.hs),
    > 537 (Semantics.thy) The Haskell function is implemented using
    > foldr, while the Isabelle function uses explicit recursion,
    > making a oneto-one comparison less obvious. If there is a
    > reason for this discrepancy, such as foldr yielding some
    > optimizations, this should be outlined in a comment.
```

@@

```console
```

### 2.6.11 Missing argument of `computeTransaction` *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Section `Relationship between Marlowe Validator and Semantics`** 
> 
> The specification mentions the output datum as the (fifth) argument for the `computeTransaction` function, while it is not an argument to it.

TODO: @bwbush will add text here.

```console
commit 00a0c2406665f06753aecf6dbb63933edb275436
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 12:39:22 2023 -0700

    PLT-4169 Fixed correspondence between semantics and validator.

    This addresses the audit comment:

    > File marlowe-cardano-specification.md, Section Relationship
    > between Marlowe Validator and Semantics The specification
    > mentions the output datum as the (fifth) argument for the
    > computeTransaction function, while it is not an argument to it.
```

@@

### 2.6.12 Missing `smallMarloweValidator` *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Various sections ``** 
> 
> The specification mentions `smallMarloweValidator` in a few places, but it is never mentioned in the source code.

TODO: @bwbush will add text here.

```console
commit 87cf548d9ae7934a702ea1a1c85d7d9bd4993aa9
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 12:37:21 2023 -0700

    PLT-4169 Fixed references to `smallMarloweValidator`.

    This addresses the audit comment:

    > File marlowe-cardano-specification.md, Various sections
    > The specification mentions smallMarloweValidator in a few
    > places, but it is never mentioned in the source code.
```

@@

### 2.6.13 Incorrect constraint reference *(Severity: Low)*

> **File `Scripts.hs`, Function `mkRolePayoutValidator`, line *150***
> 
> This line should refer to Constraint 17 rather than Constraint 16.

TODO: @bwbush will add text here.

```console
commit 3fac0d1749d107f13743756aa50c9723e776e391
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 12:42:01 2023 -0700

    PLT-4169 Fixed incorrect constraint reference.

    This addresses the audit comment:

    > File Scripts.hs, Function mkRolePayoutValidator, line 150 This line should refer to Constraint 17
    rather than Constraint 16.
```

@@

### 2.6.14 `MarloweParams` differs from the specification *(Severity: Low)*

> **File `Semantics.hs`, type `MarloweParams`, line *355***
> 
> The specification defines `MarloweParams` to contain just the payout validator hash, while the definition in the Haskell code contains just the roles currency symbol.

TODO: @bwbush will add text here.

```console
commit 8691f3351e55758cc750eee0800b1458f0a8c554
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 12:44:53 2023 -0700

    PLT-4169 Fixed `MarloweParams` in specification.

    This addresses the audit comment:

    > File Semantics.hs, type MarloweParams, line 355 The specification
    > defines MarloweParams to contain just the payout validator hash,
    > while the definition in the Haskell code contains just the roles
    > currency symbol.
```

@@

TODO: @bwbush will add text here.

```console
```

### 2.7.2 Missing tests *(Severity: Medium)*

> **File `Spec/Marlowe/Semantics/Compute.hs`, ``** 
> 
> The following properties could additionally be tested for `computeTransaction`:
> -   payment subtracts from an account,
> -   deposit adds to an account,
> -   `INotify` input produces the expected continuation,
> -   `IChoice` input produces the expected continuation,
> -   the hash of a successfully applied merkleized input matches the hash of the merkleized case.
> 
> Some of these are tested in `Spec/Marlowe/Semantics/Functions.hs` already for auxiliary functions.

TODO: @bwbush will add text here.

```console
commit ee2e52220e25736af29a1f01144d5fafbe9eed76
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 07:36:59 2023 -0700

    PLT-4168 Tests that choice and deposit continue as expected.

commit 333eee293aa3afe95d0aa7f678c7e7b7bad660ea
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 07:18:27 2023 -0700

    PLT-4168 Test that choice produces expected continuation.

commit 477284c78095482eba63adc2e6090ea68553db9d
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 06:54:14 2023 -0700

    PLT-4168 Tested that deposits add to account.

commit d241669bfb7ec8dad1cc92a7ca6185d08229817e
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Mon Mar 6 14:38:09 2023 -0700

    PLT-4168 Tested that payout substracts from account.

commit f33fc7206522031c93253323f3b8ca083d8ca8c7
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Tue Mar 7 09:56:30 2023 -0700

    PLT-4168 Tested that Merkleization continues as expected.
```

> **File `Spec/Marlowe/Semantics/Functions.hs`, `Missing merkleization tests`** 
> 
> The properties in this module do not seem to be tested with merkleized contracts or inputs except for `checkGetContinuation`. More merkleization tests should be added.

TODO: @bwbush will add text here.

```console
commit 2a936bf1bbb890cc811b966b46a10fd38314dc25
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Mon Mar 6 11:40:56 2023 -0700

    PLT-4168 Increased coverage of merkleization in validator tests.

    This addresses the following audit finding:

    > File Spec/Marlowe/Semantics/Functions.hs, Missing
    > merkleization tests The properties in this module
    > do not seem to be tested with merkleized contracts
    > or inputs except for checkGetContinuation, and
    > they probably should.

commit 8f8183d7f14c6bdc3841a1999077ef01ebac4144
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Mon Mar 6 11:40:56 2023 -0700

    PLT-4168 Increased coverage of merkleization in validator tests.

    This addresses the following audit finding:

    > File Spec/Marlowe/Semantics/Functions.hs, Missing
    > merkleization tests The properties in this module
    > do not seem to be tested with merkleized contracts
    > or inputs except for checkGetContinuation, and
    > they probably should.
```

> **File `Spec/Marlowe/Semantics/Compute.hs`, function `checkFixInterval`, lines *100-102***
> 
> The test `checkFixInterval` is never instantiated with an invalid interval that is in the past, meaning the function `fixInterval` is never tested for that case.

TODO: @bwbush will add text here.

```console
commit 811f048ba93db10972641c0c235864db1d6014e7
Author: Brian W Bush <brian.bush@iohk.io>
Date:   Mon Mar 6 10:16:08 2023 -0700

    PLT-4168 Added test for invalid interval in past.

    This addresses the following audit finding:

### 2023-03-27 response-to-audit-report.md

    > File Spec/Marlowe/Semantics/Compute.hs, function checkFixInterval,
    > lines 100-102 The test checkFixInterval is never instantiated
    > with an invalid interval that is in the past, meaning the function
    > fixInterval is never tested for that case.
```

### 2023-03-29 response-to-audit-report.md

# DRAFT Response to Audit Report

TODO:

- [ ] Add prefatory text and executive summary.
- [ ] Add responses and mitigation commits to Isabelle-related findings.
- [x] Add responses and mitigation commits to Haskell-related findings.
- [ ] Fix display of equations or replace ones that cannot be fixed with ellipses ("...").
- [ ] Replace broken links and cross references, or use elipses ("...").
- [ ] Verify the section numbering is correct.
- [ ] Add table of contents.
- [ ] Proofread.

## Executive Summary

TODO: @bwbush will add text here after everything else is written.

## Acknowledgement

TODO: @bwbush will add text thanking the audit team.

## Overview of Isabelle Changes

## Overview of Changes to Marlowe-Cardano Specification and Haskell (Plutus) Implementation

The audit report's main and high-severity concerns for the Plutus validators for Marlowe on Cardano fell into to four clusters:

1. Incorrect handling of negative deposits
    - "2.1.1 Negative deposits allow stealing funds"
2. Incomplete prevention of double satisfaction
    - "2.1.2 Contracts vulnerable to double satisfaction attacks"
3. Lax enforcement of invariants in Marlowe state
    - "2.1.7 Positive balances are not checked for the output state"
    - "2.1.8 Non-validated Marlowe states"
    - "2.1.9 Total balance of ending state uncomputed"
    - "2.1.10 Unchecked ending balance"
4. Difference semantics of association lists in Isabelle and Plutus
    - "2.1.12 Different insertion functions used in Isabelle and Haskell code"

Incorrect handling of negative deposits (item 1 above) was mitigated by the addition of a guard against this in Marlowe's Plutus validator so that its behavior was consistent with Marlowe semantics. Property-based tests were also added to check the correctness of the mitigation.

Although the Marlowe validator had prevented double-satisfaction among multiple copies of the Marlowe validator script running in the same transaction, it did not prevent double-satisfaction in cases where the Marlowe validator run alongside another Plutus validator in the same transaction (item 2 above). Double satisfaction has now been prevented by enforcing that the Marlowe validator is the only Plutus script that runs in transactions that make payments to parties. This allows Marlowe contracts to coordinate with other Plutus scripts, but only under conditions where double satisfaction is impossible. Once again, property-based tests were added to check the correctness of this mitigation.

The Marlowe validator had made optimistic assumptions about its own correct operation and hence did not check certain invariants in order to save on Plutus execution costs (item 3 above). The Marlowe semantics validator has now been thoroughly hardened against corruption of initial or final state so that it ensures that the three state-invariants of positive accounts, non-duplication of state entries (accounts, choices, and bound values), and a total internal value matching the script UTxO's value. This hardening was completed without unduly increasing the size (in bytes) of the validator or its Plutus execution cost: such was verified with simulated and on-chain measurement of execution costs for a library of real-life Marlowe contracts. Notably, the enforcement of the final invariants protects against funds becoming permanently locked in a Marlowe contract due to a garbled state: even in cases where the implementation of Marlowe semantics or parts of the Plutus validator are flawed, at least some of the execution paths of a Marlowe contract remain viable, so funds could be released even under nominally impossible but catastrophic difficulties. The property-based test suite was significantly enhanced to

Mitigation of the difference between association lists (item 4 above) in Isabelle (`MList`) and Plutus (`AssocMap`) was handled by the aforementioned enforcement of invariants, by line-by-line code inspection and annotation, and by thoroughly enhanced property-based testing. Work is in progress on a formal proof of the equivalence of `MList` and `AssocMap` under pre-conditions that hold within Marlowe semantics. Note that porting Isabelle's `MList` implementation to Plutus would have enlarged the Marlowe semantics validator by at least 2000 bytes and rendered it too costly to execute within the Cardano ledger's present execution-cost limits.

The audit report for Marlowe on Cardano was based on the commit hash [523f3d56](https://github.com/input-output-hk/marlowe-cardano/commit/523f3d56f22bf992ddb0b0c8a52bb7a5a188f9e9). The revisions and mitigations discussed here apply to [69468d66](https://github.com/input-output-hk/marlowe-cardano/commit/69468d6623e24a4ccd6659e378ae012c38ca01ce) of that repository. The appendices to this report list the differences between pre- and post-audit specification and validator code for Marlowe on Cardano.

The medium-severity concerns in the audit report center around missing tests or name shadowing. With the exception of "2.7.1 More precise failure checks", these have all been remedied either by code changes (in the case of the shadowing) or by the addition of a significant augmentation of the property-based test suite. The audit's low-severity concerns generally relate to insufficiently detailed text in the Marlowe Cardano specification, the need for more elaborate comments in the code, naming of variable bindings, or typographical errors. Nearly all of these finds have been addressed, with only the exception of a few cosmetic recommendations that were not adopted.

# 2 Responses to Specific Findings

@@

## 2.1 Main concerns

### 2.1.1 Negative deposits allow stealing funds *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Constraint 6`**
>
> The income from deposits is computed by adding up the deposit inputs, regardless of whether they are negative, while the semantics considers them as zero deposits. Combined with the absence of a balance check on the ending Marlowe state, this allows the ending balance to differ from the value paid to the Marlowe validator.
> 
> This disagreement can be exploited to steal money from a flawed Marlowe contract that allows a negative deposit. The issue is demonstrated in […].

This vulnerability was mitigated in commit [68791fac](https://github.com/input-output-hk/marlowe-cardano/commit/68791facc195068717cbc6e55d1e4fdbe1f4a521) by adding a guard against negative deposits in the Marlowe semantics validator. That guard ensures that the validator's semantics for negative deposits matches Marlowe's Isabelle semantics: namely a deposit of a negative quantity is treated as a deposit of zero. Thus a negative deposit will not decrement any of the account balances in the Plutus datum and the total of the internal balances will match the value held in the UTxO output to the Marlowe semantics script address.

A new golden test for negative deposits was added in commit [bbb9b8fc](https://github.com/input-output-hk/marlowe-cardano/commit/bbb9b8fc21ad93b7bcdb3b60be0f2f9c921f6bec) and property-based tests for negative deposits were added in commits [bbb9b8fc](https://github.com/input-output-hk/marlowe-cardano/commit/bbb9b8fc21ad93b7bcdb3b60be0f2f9c921f6bec), [9d5443c5](https://github.com/input-output-hk/marlowe-cardano/commit/9d5443c5e9306eaf49ac5389df455e07555c9e04), [4b81ee96](https://github.com/input-output-hk/marlowe-cardano/commit/4b81ee96b1e7c74c4964cf25f6bb599c80d07a52), and [15446073](https://github.com/input-output-hk/marlowe-cardano/commit/15446073131c4098fed781cdcc701b89b77f6bd0).

@@

### 2.1.2 Contracts vulnerable to double satisfaction attacks *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Constraint 15`**
> 
> No datum is required for outputs fulfilling payments to addresses generated by the evaluation of a Marlowe contract. This implies that these outputs are vulnerable to double satisfaction in transactions involving other contracts that pay to the same wallets. An example is discussed in […].
> 
> One way to strengthen the implementation is for the Marlowe validator to demand that outputs paid to addresses contain a datum that identifies the contract instance, like the `TxOutRef` of the validator UTxO being spent. Then cooperation with other contracts is possible without double satisfaction if the validators of the other contracts demand a different datum for their outputs.

The Marlowe semantics validator had prevented a limited form of double-satisfaction attack by preventing two Marlowe validators to execute in the same transaction. The audit report correctly highlights that such an attack could be feasible in cases where another Plutus validator (not the same version of the Marlowe semantics validator) executes in a transactions where a payment to an address or to the Marlowe role-payout validator satisfies both the Marlowe semantics validator and the other validator.

Although the audit report's suggestion of including a unique datum (such as that identifying the contract instance) in any output payment from the contract would likely prevent double satisfaction, implementing such a change in the Marlowe validators would have entailed such extensive changes to the specification and implementation that the overall applicability of the audit findings might subsequently be called into question. Instead, a minimal change to the specification and implementation mitigates this vulnerability: the specification was amended with a "Constraint 20. Single satisfaction" that requires the Marlowe semantics validator be the only Plutus script running in the transaction *if payments to parties are being made in the transaction*. Thus, only allowing one script to validate in case of external payments completely eliminates the possibility that one payment would satisfy two scripts. Other contracts are permitted to cooperate with Marlowe in cases such as deposits, choices, and notifications where Marlowe is not making payments.

Commits [4adf115d](https://github.com/input-output-hk/marlowe-cardano/commit/4adf115dbf795e4b014eb6ccab0acb20a73e74ed) and [5f673c47](https://github.com/input-output-hk/marlowe-cardano/commit/5f673c477ccc9b38626dca505a399962b22c2675) augment the Marlowe-Cardano specification and the implementation of the Marlowe semantics validator. Commits [3f222bc3](https://github.com/input-output-hk/marlowe-cardano/commit/3f222bc363fcf3e9cd2fa0cb5206933ef597a619), [6f6331b4](https://github.com/input-output-hk/marlowe-cardano/commit/6f6331b463e7f2fddce7ec0f64b37b030f2f795b), [39fcc8aa](https://github.com/input-output-hk/marlowe-cardano/commit/39fcc8aa9d61ed97dfa56e7ce078bf5e4c4cd6cc), [435ac680](https://github.com/input-output-hk/marlowe-cardano/commit/435ac680eb8119db1c43ec968cf8ea4ef08158b6), and [9f4fc9cd](https://github.com/input-output-hk/marlowe-cardano/commit/9f4fc9cde581b6dbbb939ec9d8e6303b8b67350e) implement property-based tests for the new single-satisfaction constraint.

@@

### 2.1.3 Missing constructor in equality instance *(Severity: High)*

> **File `Semantics.hs`, Class instance `Eq ReduceWarning`, line *845***
> 
> The constructor `ReduceAssertionFailed` is not mentioned and compares `False` against itself. This might cause validators to fail checking the presence of this particular warning.

Although this `Eq` instance is not used by the Marlowe validators (so it does not have implications for validator security), the equality test has been fixed because it is a liability for off-chain code and for potential future versions of validators that might perform such an equality test. Commit [84d65a70](https://github.com/input-output-hk/marlowe-cardano/commit/84d65a70c29320fab501ad759bbc71855d1b63fc) fixes this and commit [04805a39](https://github.com/input-output-hk/marlowe-cardano/commit/04805a393c0815fe0f2a1a6d7f76d9867ffe0c14) hardens all of the other `Eq` instances to prevent similar problems re-arising if sum types are augmented in the future.

@@

> **File `specification-v3-rc1.pdf`, `Merkleization`**
> 
> There is no property about merkleization, but merkleization is implemented in the Cardano integration.
> 
> Some relevant properties could be:
> 
> 1.  The merkleized contract produces the same payments as the analogous regular contract.
> 2.  If a merkleized case input is applied successfully, it implies that the contract hash in the input corresponds to the continuation of the contract.
> 3.  Merkleizing and unmerkleizing a contract gives back the original contract.

### 2.1.7 Positive balances are not checked for the output state *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Constraint 13`** 
> 
> *Positive balances* are only checked for the input, not for the output Marlowe state. If the semantics are flawed, a transaction can produce an unspendable output that does not satisfy this constraint.
> 
> If such a transaction is accepted, no further evaluation will be possible since all subsequent transactions will be rejected due to the very same Constraint 13. This is an hypothetical attack vector, where a malicious actor could send a transaction to block a contract.

The Plutus code for the Marlowe semantics validator had been implemented with the optimistic assumption that semantics would not be flawed and that the additional execution cost of checking the output state was not warranted. However, as the audit report points out, such a flaw would result in the contract being forever blocked from further execution. Adding a check on the final output's validity would prevent a contract from ever reaching a fully blocked state. Even if the semantics or Plutus code were flawed, at least some execution paths in the contract would remain viable, so that the contract could eventually terminate and would not permanently lock funds.

Commits [0a890845](https://github.com/input-output-hk/marlowe-cardano/commit/0a890845c44ccfe4df97e765e9b9eb743ce7d580), [8855feae](https://github.com/input-output-hk/marlowe-cardano/commit/8855feae473882b82643db792327423152e45b5c), [26f024e8](https://github.com/input-output-hk/marlowe-cardano/commit/26f024e80cea0dd4ebf8bff0f1a10b97aea87894), and [201c5df9](https://github.com/input-output-hk/marlowe-cardano/commit/201c5df9de29b3b142be5ff7d0a9b56d1b378204) augment the Marlowe Cardano specification to require positive balances upon output and implements the corresponding check in the Marlowe Semantics validator.  Commit [d514bcd0](https://github.com/input-output-hk/marlowe-cardano/commit/d514bcd075732b07e1c6a73e2e3c68afa01acf2c), [7f562545](https://github.com/input-output-hk/marlowe-cardano/commit/7f562545d013dd17472b692b550ffdc7f8a383f3), and [ebab31d9](https://github.com/input-output-hk/marlowe-cardano/commit/ebab31d9ba6dd322d374b34cad3995050a0e476e) adds a property-based test for this.

@@

### 2.1.8 Non-validated Marlowe states *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Missing constraint`** 
> 
> The validator is not specified to check that the Marlowe states in the input and output datums are valid. This condition is necessary for the lemmas about the Marlowe semantics to be applicable. The Marlowe state could become invalid if there is a flaw in the implementation of the semantics.
> 
> It also could be possible for the Marlowe state to be invalid if someone pays an output to the Marlowe validator with an invalid Marlowe state. Though this problem could be addressed with off-chain checks that prevent sending transactions that spend outputs with invalid Marlowe states. If off-chain checks are used, a note in the specification about how this is handled would be helpful.
> 
> An example showing betrayed user expectations is discussed in […].
> 
> For a valid Marlowe state, the association lists for bound values, accounts, and choices have keys sorted and without duplicates.

Prompted on this finding in the audit report, the Marlowe semantics validator has been altered to enforce all invariants (positive accounts; non-duplication of accounts, choices, and bound values; a total internal account balance that exactly matches the value in the script's UTxO) on both the initial and final states. This prevents execution of a Marlowe contract that was inadvertently or purposefully created by off-chain code with an invalid initial state. It also prevents transactions that corrupt the final state; such corruption could only happen if there were a flaw in the Plutus validator or Marlowe semantics implementation, in which case that execution-path of the contract would be blocked. The [Marlowe Best Practices Guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/best-practices.md) and [Marlowe Security Guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/security.md) warn of the importance of off-chain code creating valid initial state for Marlowe contracts.

Commits [0a890845](https://github.com/input-output-hk/marlowe-cardano/commit/0a890845c44ccfe4df97e765e9b9eb743ce7d580), [8855feae](https://github.com/input-output-hk/marlowe-cardano/commit/8855feae473882b82643db792327423152e45b5c), [26f024e8](https://github.com/input-output-hk/marlowe-cardano/commit/26f024e80cea0dd4ebf8bff0f1a10b97aea87894), and [201c5df9](https://github.com/input-output-hk/marlowe-cardano/commit/201c5df9de29b3b142be5ff7d0a9b56d1b378204) enforce that both the initial and final states in the respective incoming and outgoing datum obey the aforementioned three invariants of total value, non-duplication, and positivity. That commit also adds "Constraint 19. No duplicates" to the Marlowe Cardano specification. Property-based tests in commit [d514bcd0](https://github.com/input-output-hk/marlowe-cardano/commit/d514bcd075732b07e1c6a73e2e3c68afa01acf2c), [7f562545](https://github.com/input-output-hk/marlowe-cardano/commit/7f562545d013dd17472b692b550ffdc7f8a383f3), and [ebab31d9](https://github.com/input-output-hk/marlowe-cardano/commit/ebab31d9ba6dd322d374b34cad3995050a0e476e) check the implementation.

@@

### 2.1.9 Total balance of ending state uncomputed *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Constraint 6`** 
> 
> The constraint says
> 
> > The beginning balance plus the deposits equals the ending balance plus the payments.
> 
> However, the Marlowe validator never computes the total balance of the accounts in the ending Marlowe state. Instead, the ending balance is assumed to be whatever value is paid by the transaction to the Marlowe validator. The natural language should describe precisely what is being checked.

The mitigation of the previous audit-report finding ("2.1.8 Non-validated Marlowe states") also mitigates this finding: in the validator, `checkOwnOutputConstraint marloweData finalBalance` ensures that the computed final balance of accounts matches the value output to the script's UTxO and `checkState "o" finalBalance txOutState` ensures that it matches the sum of value in the internal accounts.

Again as previously, the relevant commits are [0a890845](https://github.com/input-output-hk/marlowe-cardano/commit/0a890845c44ccfe4df97e765e9b9eb743ce7d580), [8855feae](https://github.com/input-output-hk/marlowe-cardano/commit/8855feae473882b82643db792327423152e45b5c), [26f024e8](https://github.com/input-output-hk/marlowe-cardano/commit/26f024e80cea0dd4ebf8bff0f1a10b97aea87894), and [201c5df9](https://github.com/input-output-hk/marlowe-cardano/commit/201c5df9de29b3b142be5ff7d0a9b56d1b378204) for specification and implementation and [d514bcd0](https://github.com/input-output-hk/marlowe-cardano/commit/d514bcd075732b07e1c6a73e2e3c68afa01acf2c), [7f562545](https://github.com/input-output-hk/marlowe-cardano/commit/7f562545d013dd17472b692b550ffdc7f8a383f3), and [ebab31d9](https://github.com/input-output-hk/marlowe-cardano/commit/ebab31d9ba6dd322d374b34cad3995050a0e476e) for property-based tests.

@@

### 2.1.10 Unchecked ending balance *(Severity: High)*
